---
title: "p8105_hw1_sp4170.rmd"
author: "Shihui Peng"
date: "2023-09-24"
output: github_document
---

```{r, echo = FALSE, message=FALSE}
library(moderndive)
library(tidyverse)
```

# Problem 1

I load the moderndive library, and load the early_january_weather dataset:

```{r, echo=FALSE, message=FALSE}
data("early_january_weather")
```

## Description for dataset
* The data set early_january_weather includes several variables.
  * Their names are **`r names(early_january_weather)`**.
  * There are **`r nrow(early_january_weather)` rows** and **`r ncol(early_january_weather)` columns**.
  * Within the obvervation range of this dataset, the mean temperature is **`r mean(early_january_weather$temp)` Fahrenheit**.

## Scatterplot and description

### Scatterplot

I make a scatterplot of **temp (y)** vs **time_hour (x)**; For **humid variable**, I use color points.

```{r scatterplot for time_hour & temp}
plot_hw1 = ggplot(data=early_january_weather, aes(x = time_hour, y = temp, color = humid))+ geom_point()
plot_hw1
```

### Description

Based on the scatterplot, we can find that within the database observation range, the temperature generally shows an upward trend. The daily temperature fluctuates. The humidity shows an upward trend, too.

### Export my scatterplot

Then, I export my scatterplot to my project directory and name it "scatterplot_hw1.pdf" in a format of PDF.

```{r}
ggsave("scatterplot_hw1.pdf",plot_hw1)
```


# Problem 2

## Create a data frame

* I create a data frame comprised of:

  * a random sample of size 10 from a standard Normal distribution
  * a logical vector indicating whether elements of the sample are greater than 0
  * a character vector of length 10
  * a factor vector of length 10, with 3 different factor “levels”

```{r}
hw1_df = 
  tibble(
    ran_sam = rnorm(10),
    log_vec = ran_sam > 0,
    char_vec = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
    fac_vec = factor(sample(c("level1", "level2", "level3"), size = 10, replace = TRUE))
  )
hw1_df
```

## Try to take the means
I try to take the mean of each variable in your dataframe by pulling those variables out of my data frame hw1_df. Tidyverse library has been loading from the very beginning so no need to load it again here.

```{r}
mean_ran_sam = hw1_df %>% pull(ran_sam) %>% mean()
mean_log_vec = hw1_df %>% pull(log_vec) %>% mean()
mean_char_vec = hw1_df %>% pull(char_vec) %>% mean()
mean_fac_vec = hw1_df %>% pull(fac_vec) %>% mean()
mean_ran_sam
mean_log_vec
mean_char_vec
mean_fac_vec
```

When trying to take the mean of each variable in my dataframe using pull function, the mean for my **random sample** and **logical vector** work, but the mean for my **character vector** and **factor vector** don't work.

## Convert variables

```{r}
log_num = as.numeric(hw1_df$log_vec)
char_num = as.numeric(hw1_df$char_vec)
fac_num = as.numeric(hw1_df$fac_vec)
```

* when using, we try to convert logical, character, and factor variables to numeric variables.
  * **Logical values** are coerced to numeric values, where TRUE becomes 1 and FALSE becomes 0. Therefore, when using pull to calculate the mean of our logical data, we can get a result. So this explain what has happened above when try to get the mean of logical vectors set.
  * r tries to convert **character values** into numeric, however, it can only get not available output (NA). Therefore, when using pull to calculate the mean of our character data, we fail to get a result. This can explain what has happened above when try to get the mean of our character vectors set.
  * although the **factor vector** set is regarded as integer, it has its meaning of level. Therefore, when converting, we can get numeric vectors containing the integers representing the levels. However, when we use pull function, we cannot get the mean of our factor vectors set, but we can get the mean of this set when we use as.numeric function. This might be because the actual factor levels would not be directly converted to their original integer representations. They would be converted to numeric vectors with integers corresponding to the levels. So when we simply uss pull function to calculate the mean of a factor vector, we would get NA because we are calculates the mean of the underlying integers representing the factor levels, not the mean of the original factor levels themselves. This can explain why we cannot get the mean when using pull function but can get the mean when we convert the factor vectors into numeric ones.
